Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-06-08 doc/thesis/discussion.Rnw
Up to ABC in intro now

Our implementation of \gls{SMC} uses a simple multinomial scheme to sample
particles from the population according to their weights. Several other
sampling strategies have been developed~\autocite{douc2005comparison}, and it
is possible that the use of a more sophisticated technique might increase the
algorithm's accuracy. Finally, the \gls{ABC}-\gls{SMC} algorithm is
computationally intensive, taking about a day when run on 20 cores in parallel
with the settings described in the methods section. Implementing
parallelization using MPI, rather than POSIX threads as we have done here,
would allow the program to be run over a larger number of cores on multiple
CPUs in parallel.

@@

Both \textcite{de2007preferential} and \textcite{novitsky2014impact} studied
populations whose primary risk factor for \gls{HIV} infection was heterosexual
contact. \citeauthor{de2007preferential} explicitly excluded reported
homosexual contacts; \citeauthor{novitsky2014impact} did not, but noted that
heterosexual contact is the primary mode of transmission in Botswana where the
study was done. In the first of the two papers describing the Botswana
study~\autocite{novitsky2013phylogenetic}, the authors noted that their sample
was gender-biased, being composed of approximately 75\% women. Our estimate of
\gls{alpha} for these data was X or X, depending on the prior on \gls{m};
\citeauthor{de2007preferential} estimated 0.54, 0.57, and 0.29 for 3-year,
5-year, and lifetime partnership networks respectively for the female portion
of their sample.

@@

>>>>

Our implementation of \gls{SMC} uses a simple multinomial scheme to sample
particles from the population according to their weights. Several other
sampling strategies have been developed~\autocite{douc2005comparison}, and it
is possible that the use of a more sophisticated technique might increase the
algorithm's accuracy. Finally, the \gls{ABC}-\gls{SMC} algorithm is
computationally intensive, taking about a day when run on 20 cores in parallel
with the settings described in the methods section. Implementing
parallelization using \gls{MPI}, rather than \gls{POSIX} threads as we have
done here, would allow the program to be run over a larger number of cores on
multiple CPUs in parallel.

@@

Both \textcite{de2007preferential} and \textcite{novitsky2014impact} studied
populations whose primary risk factor for \gls{HIV} infection was heterosexual
contact. \citeauthor{de2007preferential} explicitly excluded reported
homosexual contacts; \citeauthor{novitsky2014impact} did not, but noted that
heterosexual contact is the primary mode of transmission in Botswana where the
study was done. In the first of the two papers describing the Botswana
study~\autocite{novitsky2013phylogenetic}, the authors noted that their sample
was gender-biased, being composed of approximately 75\% women. Our estimate of
\gls{alpha} for these data was 
    \Sexpr{d[m_min == 1 & dataset == "novitsky2014" & parameter == "alpha", round(mean, 2)]}
        or 
    \Sexpr{d[m_min == 2 & dataset == "novitsky2014" & parameter == "alpha", round(mean, 2)]},
depending on the prior on \gls{m}; \citeauthor{de2007preferential} estimated
0.54, 0.57, and 0.29 for 3-year, 5-year, and lifetime partnership networks
respectively for the female portion of their sample.

@@

One somewhat surprising result was the difference between parameter estimates 
for the \textcite{li2015hiv} and \textcite{wang2015targeting} datasets. Both
groups studied cohorts of acutely infected \gls{MSM} in major Chinese cities
(Shanghai and Beijing respectively); yet, the \citeauthor{li2015hiv} data was 
estimated to have a lower \acrlong{PA} power and larger infected population
than the \textcite{wang2015targeting} data\ldots

% for discussion
In order to compare our results to existing literature on networks and
distributions of partner counts, we have reported estimated values for the
power law exponent \gls{gamma} of the real data sets we evaluated. However, the
posterior means for \gls{alpha} for all six datasets were less than one; the
degree distributions in this parameter range are stretched exponential, not
power law~\autocite{krapivsky2000connectivity}. As we show in
\cref{fig:powerlaw}, the power law fit does capture the slope of the degree
distribution fairly well, but the results should still be interpreted
cautiously. \textcite{krapivsky2000connectivity} showed that the power law
distribution can be maintained, with \gls{gamma} tuned to any desired value, by
a straightforward modification of the \gls{BA} model. The authors define the
``connection kernel'' $A_k$ the probability of a new connection to a node of
degree $k$, up to a normalizing constant. In the \gls{BA} model as we have
presented it here, $A_k = k^{\alpha} + 1$. Taking $A_k$ as any asymptotically
linear function will result in a power law distribution, with the exponent
\gls{gamma} determined by the properties of $A_k$. Implementing such a model
would be straightforward and seems a natural next step toward improving the
realism of the \gls{BA} model.

\textcite{rothenberg2007large} estimated the power law exponents for 15 major
network studies of sexual and injection drug use networks carried out between
1981 and 2000. The estimated \gls{gamma} values for ``all contacts'' (that is,
including both interviewed and non-interviewed individuals) were 
between 2.08 and 2.87 for all but one of the networks\ldots

### 2016-06-08 doc/thesis/discussion.tex
Up to ABC in intro now

Both \textcite{de2007preferential} and \textcite{novitsky2014impact} studied
populations whose primary risk factor for \gls{HIV} infection was heterosexual
contact. \citeauthor{de2007preferential} explicitly excluded reported
homosexual contacts; \citeauthor{novitsky2014impact} did not, but noted that
heterosexual contact is the primary mode of transmission in Botswana where the
study was done. In the first of the two papers describing the Botswana
study~\autocite{novitsky2013phylogenetic}, the authors noted that their sample
was gender-biased, being composed of approximately 75\% women. Our estimate of
\gls{alpha} for these data was X or X, depending on the prior on \gls{m};
\citeauthor{de2007preferential} estimated 0.54, 0.57, and 0.29 for 3-year,
5-year, and lifetime partnership networks respectively for the female portion
of their sample.

>>>>

Both \textcite{de2007preferential} and \textcite{novitsky2014impact} studied
populations whose primary risk factor for \gls{HIV} infection was heterosexual
contact. \citeauthor{de2007preferential} explicitly excluded reported
homosexual contacts; \citeauthor{novitsky2014impact} did not, but noted that
heterosexual contact is the primary mode of transmission in Botswana where the
study was done. In the first of the two papers describing the Botswana
study~\autocite{novitsky2013phylogenetic}, the authors noted that their sample
was gender-biased, being composed of approximately 75\% women. Our estimate of
\gls{alpha} for these data was 
    0.55
        or 
    0.53,
depending on the prior on \gls{m}; \citeauthor{de2007preferential} estimated
0.54, 0.57, and 0.29 for 3-year, 5-year, and lifetime partnership networks
respectively for the female portion of their sample.

### 2016-06-08 doc/thesis/intro.tex
Up to ABC in intro now

One mechanism which has been shown to lead to scale-free networks is
\defn{preferential attachment}~\autocite{simon1955class,
barabasi1999emergence}. The simplest preferential attachment model is known as
the \gls{BA} model after its inventors~\autocite{barabasi1999emergence}. Under
this model, networks are formed by starting with a small number $m_0$ of nodes.
New nodes are added one at a time until there are a total of $N$ in the
network. Each time a new node is added, $m \geq 1$ edges are added from it to
other nodes in the graph. In the original formulation, the partners of the new
node are chosen with probability linearly proportional to their degree.
However, \citeauthor{barabasi1999emergence} suggest extending the model such
that the probability of choosing a partner of degree $d$ is proportional to
$d^\alpha + 1$ for some constant $\alpha$, and we use this extension here. When
$m = 1$, the network takes on the distinctive shape of a tree, that is, it does
not contain any cycles. Cycles are present in the network for all other $m$
values. Examples of \gls{BA} networks with three different values of the
preferential attachment power \gls{alpha} are shown in \cref{fig:baeg}.

@@

There has been some contention over the idea that contact networks are
scale-free. \textcite{handcock2004likelihood} fit several stochastic models of
partner formation to empirical degree distributions derived from population
surveys of sexual behaviour. They found that a negative binomial distribution,
rather than a power law, was the best fit to five out of six datasets, although
the difference in goodness of fit was extremely small in four out of these
five. \textcite{bansal2007individual} found that an exponential distribution,
rather than a power law, was the best fit to degree distributions of six social
and sexual networks.

@@

{\color{blue}\uline{
Recall that the primary objective of our work is to develop a statistical
inference method for estimating contact network parameters from transmission
trees. In particular, for a network model with parameters $\theta$ and an input
transmission tree $T$, we are interested in the posterior distribution}
\begin{align}
  \Pr(\theta \mid T) = \frac{\Pr(T \mid \theta) \Pr(\theta)}{\Pr(T)},
  \label{eq:post}
\end{align}
where we have adopted the common abuse of the symbol $\Pr$ to refer to a
probability density. \uline{As alluded to in} \cref{sec:obj}, \uline{both the
likelihood $\Pr(T \mid \theta)$ and the normalizing constant $\Pr(T)$ are
likely computationally intractable (this will be explained further in}
\cref{subsec:just}\uline{). Hence, rather than computing the posterior
distribution analytically, we will approximate it using a \defn{Monte Carlo}
approach. The fundamental idea behind Monte Carlo methods is succinctly
expressed by \textcite{liu2001theoretical}}:
\begin{quote}
    \uline{Monte Carlo's view of the world is that any probability distribution $\pi$, 
    regardless of its complexity, can always be \emph{represented} by a
    discrete sample from it. By ``represented'', we mean that any computation
    of expectations using $\pi$ can be replaced to an acceptable degree of
    accuracy by using the empirical distribution resulting from the discrete
    sample.}
\end{quote}
\uline{In other words, if we are able to sample enough points from a
distribution of interest, we will be able to make reasonably accurate
statements about the distribution itself. For example, the expected value of
the distribution can be estimated by the sample's population mean. The reason
Monte Carlo methods will be useful in this work is that algorithms exist for
obtaining samples from distributions that are analytically intractable and from
which direct sampling is not possible } (for a review see
\autocite{robert2004monte}). \uline{\Gls{SMC}}~\autocite{doucet2000sequential,
doucet2001introduction, liu2008monte} \uline{is one such algorithm.}

\uline{\Gls{SMC} is also known as the \emph{particle filter}. Rather than
sampling points one at a time from the target distribution, \gls{SMC} considers
a population of points or ``particles'', here denoted $\set{x^{(k)}}$. The
particles are associated with weights, $\set{w(x^{(k)})}$. These weighted
particles are a representation for the target distribution. For example, the 
expected value of the distribution is approximated by}
\[
  \frac{1}{n} \sum_{k=1}^n x^{(k)} w(x^{(k)}),
\]
\uline{where $n$ is the number of particles. Initially, the particles do not
represent the target distribution but rather a more tractable distribution from
which direct sampling is straightforward. The word ``sequential'' is used to
describe the iterative process of perturbation, resampling, and reweighting
applied to the particles in such a way that they converge, collectively, to a
sample from the target.}

@@

{\color{red}\sout{Ordinarily, Bayesian inference targets the posterior
distribution $\Pr(\theta \mid y)$. That is, in the Bayesian framework,}} 
{\color{blue}\uline{By targeting the posterior distribution, Bayesian inference
makes the assertion that}} model parameters with higher posterior density are
``better'' in the sense that they offer a more credible explanation for the
observed data. The approximate posterior targeted by \gls{ABC} uses an
alternative metric for parameter credibility, namely the similarity of
simulated datasets to the observed data. If datasets simulated under the model
closely resemble the real data, it follows that the model is a reasonable
approximation to the real-world process generating the observed data. More
formally, {\color{blue}\uline{let $y$ be the observed data to which we are
trying to fit a model with parameters $\theta$. In the case of this work, the
data is a transmission tree $T$, but we shall stick with the generic variable
$y$ for now.}} Suppose we have a distance measure $\rho$ defined on the space
of all possible data our model could generate. \gls{ABC} aims to sample from
the joint posterior distribution of model parameters and simulated datasets $z$
which are within some small distance $\varepsilon$ of the observed data $y$,
\[
  \pi_{\varepsilon}(\theta, z \mid y) =
  \frac{\pi(\theta) \Pr(z \mid \theta) \I_{A_{\varepsilon, y}} (z)}
  {\int_{A_{\varepsilon, y} \times \Theta} \pi(\theta) \Pr(z \mid \theta) \d \theta}.
\]
Here, $A_{\varepsilon, y}$ is an $\varepsilon$-ball around $y$ with
respect to $\rho$, $\Theta$ is the space of all possible model parameters, and
$\I$ is the indicator function~\autocite{marin2012approximate}.
{\color{blue}\uline{The term $\Pr(z \mid \theta)$ appears to be the bothersome
likelihood again, but this will turn out not to be a problem because we are
simulating $z$ ourselves. In fact, it is possible to sample from this
distribution exactly (see next section).}}

{\color{blue}\uline{To return to the context of this thesis, the observed data
$y$ is an estimated transmission tree for a viral epidemic under investigation.
The model in question is a contact network model with parameters $\theta$. $z$
is a simulated transmission tree, obtained by first generating a contact
network under the model, and then simulating the spread of an epidemic over
that network. A transmission tree can be constructed by keeping track of who
infected whom during the simulated epidemic (further details will be given in
} \cref{subsec:nettree} \uline{). The distance function $\rho$ must compare two
transmission trees - the observed tree $y$, and the simulated tree $z$. We will
define this distance function using the tree kernel discussed in}
\cref{subsec:treeshape}. \uline{In words, the approximate posterior we consider
here is a distribution which assigns a joint probability density to model
parameters and simulated transmission trees under those parameters. The
probability density is proportional to the product of the prior on the
parameters, and the likelihood of the simulated transmission tree under those
parameters, but only if the simulated transmission tree is sufficiently close
to the true tree. Otherwise, the probability density is zero. }}

{\color{red}\sout{As we shall see in the next section, this distribution can
be sampled from exactly. The word ``approximate'' derives from the assumption
that,}}
for a suitably chosen distance $\rho$ and a small enough $\varepsilon$, the
marginal in $z$ of this distribution approximates the posterior of
interest~\autocite{marin2012approximate}. That is,
\[
  \int \pi_\varepsilon(\theta, z \mid y) \d z \approx f(\theta \mid y).
\]
The distribution $\pi_\varepsilon(\theta, z \mid y)$ is variously referred to
as the \textit{\gls{ABC} target distribution} or the \gls{ABC} approximation
to the posterior. The intuition for why this approximation might be reasonable
comes from the fact that, when $\varepsilon = 0$, the $\varepsilon$-ball around
$y$ should contain only $y$ itself, hence the integral on the left is exactly
equal to the posterior. Thus, by taking $\varepsilon$ small, we should attain
something close to the posterior {\color{blue}\uline{if $\rho$ captures the
similarity between datasets reasonably well. However, the accuracy of the
\gls{ABC} approximation depends heavily on the choice of distance function
\autocite{aeschbacher2012novel,blum2013comparative}.}}

>>>>

One mechanism which has been shown to lead to scale-free networks is
\defn{preferential attachment}~\autocite{simon1955class,
barabasi1999emergence}. The simplest preferential attachment model is known as
the \gls{BA} model after its inventors~\autocite{barabasi1999emergence}. Under
this model, networks are formed by starting with a small number $m_0$ of nodes.
New nodes are added one at a time until there are a total of $N$ in the
network. Each time a new node is added, $m \geq 1$ edges are added from it to
other nodes in the graph. In the original formulation, the partners of the new
node are chosen with probability linearly proportional to their degree plus
one.

There has been some contention over the idea that contact networks are
scale-free. \textcite{handcock2004likelihood} fit several stochastic models of
partner formation to empirical degree distributions derived from population
surveys of sexual behaviour. They found that a negative binomial distribution,
rather than a power law, was the best fit to five out of six datasets, although
the difference in goodness of fit was extremely small in four out of these
five. \textcite{bansal2007individual} found that an exponential distribution,
rather than a power law, was the best fit to degree distributions of six social
and sexual networks. \textcite{drombowski2013} contend that sexual networks are
shaped more by homophily (``like attract like'') than by preferential
attachment, but find that \gls{IDU} network do demonstrate a scale-free
structure.

In the paper describing the \gls{BA} model, \citeauthor{barabasi1999emergence}
suggest an extension where the probability of choosing a partner of degree $d$
is proportional to $d^\alpha + 1$ for some constant \gls{alpha}.
{\color{blue}\uline{When $\gls{alpha} \neq 1$, the degree distribution no
longer follows a power law~\autocite{krapivsky2000connectivity}. For
$\gls{alpha} < 1$, the distribution is a stretched exponential, meaning that
the number of nodes of degree $k$ is proportional to $\exp(-k^\beta)$ for some
constant $\beta$. For $\gls{alpha} > 1$, the distribution takes on a
characteristic called \emph{gelation}, where a one or a few high-degree hub
nodes are connected to nearly every other node in the graph. We do not believe
these departures from the power law affect the applicability of the model to
real world networks. In fact,} \textcite{de2007preferential} \uline{were able to
estimate the \acrlong{PA} power from partner count data collected from the same
individuals for consecutive time intervals, and found a value less than one in
all cases. It is also worth noting that, in addition to the \gls{BA} model,
other investigations of the interaction between contact networks and
transmission trees have studied the \acrlong{ER} and \acrlong{WS}
models~\autocite{leventhal2012inferring}, whose degree distributions do not
generally follow a power law under any parameter settings.}}

When $m = 1$, the network takes on the distinctive shape of a tree, that is, it
does not contain any cycles. Cycles are present in the network for all other
$m$ values. Examples of \gls{BA} networks with three different values of the
preferential attachment power \gls{alpha} are shown in \cref{fig:baeg}.

@@

@@

{\color{blue}\uline{
Recall that the primary objective of our work is to develop a statistical
inference method for estimating contact network parameters from transmission
trees. In particular, for a network model with parameters $\theta$ and an input
transmission tree $T$, we are interested in the posterior distribution}
\begin{align}
  \Pr(\theta \mid T) = \frac{\Pr(T \mid \theta) \Pr(\theta)}{\Pr(T)},
  \label{eq:post}
\end{align}
\uline{where we have adopted the common abuse of the symbol $\Pr$ to refer to a
probability density. As alluded to in} \cref{sec:obj}, \uline{both the
likelihood $\Pr(T \mid \theta)$ and the normalizing constant $\Pr(T)$ are
likely computationally intractable (this will be explained further in}
\cref{subsec:just}\uline{). Hence, rather than computing the posterior
distribution analytically, we will approximate it using a \defn{Monte Carlo}
approach. The fundamental idea behind Monte Carlo methods is succinctly
expressed by \textcite{liu2001theoretical}}:
\begin{quote}
    \uline{Monte Carlo's view of the world is that any probability distribution $\pi$, 
    regardless of its complexity, can always be \emph{represented} by a
    discrete sample from it. By ``represented'', we mean that any computation
    of expectations using $\pi$ can be replaced to an acceptable degree of
    accuracy by using the empirical distribution resulting from the discrete
    sample.}
\end{quote}
\uline{In other words, if we are able to sample enough points from a
distribution of interest, we will be able to make reasonably accurate
statements about the distribution itself. For example, the expected value of
the distribution can be estimated by the sample's population mean. The reason
Monte Carlo methods will be useful in this work is that algorithms exist for
obtaining samples from distributions that are analytically intractable and from
which direct sampling is not possible } (for a review see
\autocite{robert2004monte}). \uline{\Gls{SMC}}~\autocite{doucet2000sequential,
doucet2001introduction, liu2008monte} \uline{is one such algorithm.}

\uline{\Gls{SMC} is also known as the \emph{particle filter}. Rather than
sampling points one at a time from the target distribution, \gls{SMC} considers
a population of points or ``particles'', here denoted $\set{x^{(k)}}$ and
indexed by an integer $k$. The particles are associated with weights,
$\set{w(x^{(k)})}$. These weighted particles are a representation for the
target distribution. For example, the expected value of the distribution is
approximated by}
\[
  \frac{1}{n} \sum_{k=1}^n x^{(k)} w(x^{(k)}),
\]
\uline{where $n$ is the number of particles. Initially, the particles do not
represent the target distribution but rather a more tractable distribution from
which direct sampling is straightforward. The word ``sequential'' is used to
describe the iterative process of perturbation, resampling, and reweighting
applied to the particles in such a way that they converge, collectively, to a
sample from the target.}

@@

{\color{red}\sout{Ordinarily, Bayesian inference targets the posterior
distribution $\Pr(\theta \mid y)$. That is, in the Bayesian framework,}} 
{\color{blue}\uline{By targeting the posterior distribution, Bayesian inference
makes the assertion that}} model parameters with higher posterior density are
``better'' in the sense that they offer a more credible explanation for the
observed data. The approximate posterior targeted by \gls{ABC} uses an
alternative metric for parameter credibility, namely the similarity of
simulated datasets to the observed data. If datasets simulated under the model
closely resemble the real data, it follows that the model is a reasonable
approximation to the real-world process generating the observed data. More
formally, {\color{blue}\uline{let $y$ be the observed data to which we are
trying to fit a model with parameters $\theta$. In the case of this work, the
data is a transmission tree $T$, but we shall stick with the generic variable
$y$ for now.}} Suppose we have a distance measure $\rho$ defined on the space
of all possible data our model could generate. \gls{ABC} aims to sample from
the joint posterior distribution of model parameters and simulated datasets $z$
which are within some small distance $\varepsilon$ of the observed data $y$,
\[
  \pi_{\varepsilon}(\theta, z \mid y) =
  \frac{\Pr(\theta) \Pr(z \mid \theta) \I_{A_{\varepsilon, y}} (z)}
  {\int_{A_{\varepsilon, y} \times \Theta} \Pr(\theta) \Pr(z \mid \theta) \d \theta}.
\]
Here, $A_{\varepsilon, y}$ is an $\varepsilon$-ball around $y$ with
respect to $\rho$, $\Theta$ is the space of all possible model parameters, and
$\I$ is the indicator function~\autocite{marin2012approximate}.
{\color{blue}\uline{The term $\Pr(z \mid \theta)$ appears to be the bothersome
likelihood again, but this will turn out not to be a problem because we are
simulating $z$ ourselves. In fact, it is possible to sample from this
distribution exactly (see next section).}}

{\color{blue}\uline{To return to the context of this thesis, the observed data
$y$ is an estimated transmission tree for a viral epidemic under investigation.
The model in question is a contact network model with parameters $\theta$. $z$
is a simulated transmission tree, obtained by first generating a contact
network under the model, and then simulating the spread of an epidemic over
that network. A transmission tree can be constructed by keeping track of who
infected whom during the simulated epidemic (further details will be given in
} \cref{subsec:nettree} \uline{). The distance function $\rho$ must compare two
transmission trees - the observed tree $y$, and the simulated tree $z$. We will
define this distance function using the tree kernel discussed in}
\cref{subsec:treeshape}. \uline{In words, the approximate posterior we consider
here is a distribution which assigns a joint probability density to model
parameters and simulated transmission trees under those parameters. The
probability density is proportional to the product of the prior on the
parameters, and the likelihood of the simulated transmission tree under those
parameters, but only if the simulated transmission tree is sufficiently close
to the true tree. Otherwise, the probability density is zero.}}

{\color{red}\sout{As we shall see in the next section, this distribution can
be sampled from exactly. The word ``approximate'' derives from the assumption
that,}}
{\color{blue}\uline{It is not yet obvious that the \gls{ABC} target
distribution just presented can in fact be used to approximate the true
posterior. TODO }}

for a suitably chosen distance $\rho$ and a small enough $\varepsilon$, the
marginal in $z$ of this distribution approximates the posterior of
interest~\autocite{marin2012approximate}. That is,
\[
  \int \pi_\varepsilon(\theta, z \mid y) \d z \approx f(\theta \mid y).
\]
The distribution $\pi_\varepsilon(\theta, z \mid y)$ is variously referred to
as the \textit{\gls{ABC} target distribution} or the \gls{ABC} approximation
to the posterior. The intuition for why this approximation might be reasonable
comes from the fact that, when $\varepsilon = 0$, the $\varepsilon$-ball around
$y$ should contain only $y$ itself, hence the integral on the left is exactly
equal to the posterior. Thus, by taking $\varepsilon$ small, we should attain
something close to the posterior {\color{blue}\uline{if $\rho$ captures the
similarity between datasets reasonably well. However, the accuracy of the
\gls{ABC} approximation depends heavily on the choice of distance function
\autocite{aeschbacher2012novel,blum2013comparative}.}}

### 2016-06-08 doc/thesis/preamble.tex
Moar
Formatting for GPS
Underline changes for printing
Started working on Sally's revisions
More of Sally's revisions
More of Sally's changes
More exposition for ABC; rewrote discussion of real data
Up to ABC in intro now

\documentclass[11pt]{report}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage[top=1in, bottom=1in, inner=1.25in, outer=1in]{geometry}
\usepackage{setspace}
\usepackage[backend=biber, style=numeric-comp, sorting=none]{biblatex}
\usepackage{nicefrac}
\usepackage{amsmath, amssymb}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{multicol}
\usepackage{multirow}
\usepackage[hidelinks]{hyperref}
\usepackage[acronym,nonumberlist]{glossaries}
\usepackage{glossary-mcols}
\usepackage{cleveref}
\usepackage{array}
\usepackage{textcomp}
\usepackage{float}
\usepackage{pdflscape}
\usepackage{fancyhdr}
\usepackage[dvipsnames]{xcolor}

@@

\newcommand{\defn}[1]{\textit{#1}}
\newcommand{\software}[1]{\textit{#1}}
\newcommand{\etal}{\textit{et al.}}
\newcommand{\eg}{\textit{e.g.}}

@@

\newcommand{\N}{\mathcal{N}}
\newcommand{\I}{\mathbb{I}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\X}{\mathcal{X}}
\renewcommand{\L}{\mathcal{L}}

@@

\DeclareMathOperator{\Exponential}{Exponential}
\DeclareMathOperator{\Uniform}{Uniform}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\inc}{in}
\DeclareMathOperator{\out}{out}
\DeclareMathOperator{\ESS}{ESS}
\DeclareMathOperator{\nc}{nc}

@@

\newacronym{MCMC}{MCMC}{Markov chain Monte Carlo}
\newacronym{ER}{ER}{Erd\H{o}s-R\'enyi}
\newacronym{BA}{BA}{Barab\'asi-Albert}
\newacronym{WS}{WS}{Watts-Strogatz}
\newacronym{SI}{SI}{susceptible-infected}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{SVR}{SVR}{support vector regression}
\newacronym{kSVM}{kSVM}{kernel support vector machine}
\newacronym{kSVR}{kSVR}{kernel support vector regression}
\newacronym{PCA}{PCA}{principal components analysis}
\newacronym{kPCA}{kPCA}{kernel principal components analysis}
\newacronym{SMC}{SMC}{sequential Monte Carlo}
\newacronym{ABC}{ABC}{approximate Bayesian computation}
\newacronym{nltt}{nLTT}{normalized lineages-through-time}
\newacronym{ltt}{LTT}{lineages-through-time}
\newacronym{pdf}{pdf}{probability density function}
\newacronym{rv}{rv}{random variable}
\newacronym{ML}{ML}{maximum likelihood}
\newacronym{MAP}{MAP}{maximum \textit{a posteriori}}
\newacronym{GSL}{GSL}{GNU scientific library}
\newacronym{SIR}{SIR}{susceptible-infected-recovered}
\newacronym{IS}{IS}{importance sampling}
\newacronym{SIS}{SIS}{sequential importance sampling}
\newacronym{HMM}{HMM}{hidden Markov model}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{HPD}{HPD}{highest posterior density}
\newacronym{HDI}{HDI}{highest density interval}
\newacronym{HIV}{HIV}{human immunodeficiency virus}
\newacronym{SARS}{SARS}{severe acute respiratory syndrome}
\newacronym{ODE}{ODE}{ordinary differential equation}
\newacronym{GLM}{GLM}{generalized linear model}
\newacronym{tasp}{TasP}{treatment as prevention}
\newacronym{MH}{MH}{Metropolis-Hastings}
\newacronym{GTR}{GTR}{generalized time-reversible}
\newacronym{ESS}{ESS}{expected sample size}
\newacronym{MSM}{MSM}{men who have sex with men}
\newacronym{ERGM}{ERGM}{exponential random graph model}
\newacronym{IDU}{IDU}{injection drug users}

\newglossaryentry{alpha}{
  name=$\alpha$,
  description=preferential attachment power parameter in Barab\'asi-Albert networks
}
\newglossaryentry{m}{
  name=$m$,
  description=number of edges added per vertex when constructing a Barab\'asi-Albert network
}
\newglossaryentry{N}{
  name=$N$,
  description=total number of nodes in a contact network
}
\newglossaryentry{I}{
  name=$I$,
  description=number of infected nodes in a contact network at the time of
  transmission tree sampling
}
\newglossaryentry{gamma}{
  name=$\gamma$,
  description=exponent of power-law degree distribution in scale-free networks
}
\newglossaryentry{p}{
  name=$p$,
  description=edge density parameter in Erd\H{o}s-R\'enyi networks
}
\newglossaryentry{lambda}{
  name=$\lambda$,
  description=decay factor meta-parameter for tree kernel
}
\newglossaryentry{sigma}{
  name=$\sigma$,
  description=radial basis function variance meta-parameter for tree kernel
}
\newglossaryentry{n}{
  name=$n$,
  description=number of particles used for ABC-SMC
}
\newglossaryentry{M}{
  name=$M$,
  description=number of simulated datasets per particle in ABC-SMC
}
\newglossaryentry{tmrca}{
  name=$t_{\mathrm{mrca}}$,
  description=time to most recent common ancestor of a group of extant taxa
}

>>>>

\documentclass[11pt]{report}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage[top=1in, bottom=1in, inner=1.25in, outer=1in]{geometry}
\usepackage{setspace}
\usepackage[backend=biber, style=numeric-comp, sorting=none]{biblatex}
\usepackage{nicefrac}
\usepackage{amsmath, amssymb}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{multicol}
\usepackage{multirow}
\usepackage[hidelinks]{hyperref}
\usepackage[acronym,nonumberlist]{glossaries}
\usepackage{glossary-mcols}
\usepackage{cleveref}
\usepackage{array}
\usepackage{textcomp}
\usepackage{float}
\usepackage{pdflscape}
\usepackage{fancyhdr}
\usepackage[dvipsnames]{xcolor}
\usepackage[titletoc]{appendix}
\usepackage{tocloft}
\usepackage[normalem]{ulem}

@@

\newcommand{\defn}[1]{\textit{#1}}
\newcommand{\software}[1]{\textit{#1}}
\newcommand{\etal}{\textit{et al.}}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\ie}{\textit{i.e.}}

@@

\newcommand{\N}{\mathcal{N}}
\newcommand{\I}{\mathbb{I}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\E}{\mathbb{E}}
\renewcommand{\L}{\mathcal{L}}

@@

\DeclareMathOperator{\Exponential}{Exponential}
\DeclareMathOperator{\Uniform}{Uniform}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\inc}{in}
\DeclareMathOperator{\out}{out}
\DeclareMathOperator{\ESS}{ESS}
\DeclareMathOperator{\nc}{nc}
\DeclareMathOperator{\nl}{nl}

@@

\newacronym{MCMC}{MCMC}{Markov chain Monte Carlo}
\newacronym{ER}{ER}{Erd\H{o}s-R\'enyi}
\newacronym{BA}{BA}{Barab\'asi-Albert}
\newacronym{WS}{WS}{Watts-Strogatz}
\newacronym{SI}{SI}{susceptible-infected}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{SVR}{SVR}{support vector regression}
\newacronym{kSVM}{kSVM}{kernel support vector machine}
\newacronym{kSVR}{kSVR}{kernel support vector regression}
\newacronym{PCA}{PCA}{principal components analysis}
\newacronym{kPCA}{kPCA}{kernel principal components analysis}
\newacronym{SMC}{SMC}{sequential Monte Carlo}
\newacronym{ABC}{ABC}{approximate Bayesian computation}
\newacronym{nltt}{nLTT}{normalized lineages-through-time}
\newacronym{ltt}{LTT}{lineages-through-time}
\newacronym{pdf}{pdf}{probability density function}
\newacronym{rv}{rv}{random variable}
\newacronym{ML}{ML}{maximum likelihood}
\newacronym{MAP}{MAP}{maximum \textit{a posteriori}}
\newacronym{GSL}{GSL}{GNU scientific library}
\newacronym{SIR}{SIR}{susceptible-infected-recovered}
\newacronym{IS}{IS}{importance sampling}
\newacronym{SIS}{SIS}{sequential importance sampling}
\newacronym{HMM}{HMM}{hidden Markov model}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{HPD}{HPD}{highest posterior density}
\newacronym{HDI}{HDI}{highest density interval}
\newacronym{HIV}{HIV}{human immunodeficiency virus}
\newacronym{SARS}{SARS}{severe acute respiratory syndrome}
\newacronym{ODE}{ODE}{ordinary differential equation}
\newacronym{GLM}{GLM}{generalized linear model}
\newacronym{tasp}{TasP}{treatment as prevention}
\newacronym{MH}{MH}{Metropolis-Hastings}
\newacronym{GTR}{GTR}{generalized time-reversible}
\newacronym{ESS}{ESS}{expected sample size}
\newacronym{MSM}{MSM}{men who have sex with men}
\newacronym{ERGM}{ERGM}{exponential random graph model}
\newacronym{IDU}{IDU}{injection drug users}
\newacronym{ANOVA}{ANOVA}{analysis of variance}
\newacronym{MANOVA}{MANOVA}{multivariate analysis of variance}
\newacronym{AIC}{AIC}{Akaike information criterion}
\newacronym{POSIX}{POSIX}{Portable Operating System Interface}
\newacronym{IQR}{IQR}{interquartile range}
\newacronym{MPI}{MPI}{message passing interface}
\newacronym{PA}{PA}{preferential attachment}

\newglossaryentry{alpha}{
  name=\ensuremath{\alpha},
  description=preferential attachment power parameter in Barab\'asi-Albert networks
}
\newglossaryentry{m}{
  name=$m$,
  description=number of edges added per vertex when constructing a Barab\'asi-Albert network
}
\newglossaryentry{N}{
  name=$N$,
  description=total number of nodes in a contact network
}
\newglossaryentry{I}{
  name=$I$,
  description=number of infected nodes in a contact network at the time of
  transmission tree sampling
}
\newglossaryentry{gamma}{
  name=$\gamma$,
  description=exponent of power-law degree distribution in scale-free networks
}
\newglossaryentry{p}{
  name=$p$,
  description=edge density parameter in Erd\H{o}s-R\'enyi networks
}
\newglossaryentry{lambda}{
  name=$\lambda$,
  description=decay factor meta-parameter for tree kernel
}
\newglossaryentry{sigma}{
  name=$\sigma$,
  description=radial basis function variance meta-parameter for tree kernel
}
\newglossaryentry{n}{
  name=$n$,
  description=number of particles used for ABC-SMC
}
\newglossaryentry{M}{
  name=$M$,
  description=number of simulated datasets per particle in ABC-SMC
}
\newglossaryentry{tmrca}{
  name=$t_{\mathrm{mrca}}$,
  description=time to most recent common ancestor of a group of extant taxa
}
\newglossaryentry{alphaess}{
  name=$\alpha_{\ESS}$,
  description=per-iteration rate of expected sample size decay in adaptive ABC-SMC
}
